Fast Bayesian Inference for Exoplanet Discovery in Radial Velocity Data

### 2014-12-13 Paper/ms.tex
sigma_extra

The unknown parameters are:
\begin{eqnarray}
\left\{N, \boldsymbol{\alpha}, \{\boldsymbol{\psi}\}_{i=1}^N, y_0, s, \nu\right\}
\end{eqnarray}
where $N$ is the number of planets,
$\boldsymbol{\alpha} = \{\mu_P, \sigma_P, \mu_A\}$ are the
hyperparameters used to define the prior for the properties of the planets,
and $\psi_i = \{P_i, A_i, \phi_i, e_i, \theta_i\}$
are the properties of planet $i$.
The parameter $y_0$ describes a DC offset in the data, and $s$ and $\nu$ are
parameters allowing for a non-gaussian distribution for the noise in the
measurements.

A standard assumption for the probability distribution of the data given the
parameters (known as the sampling distribution, which becomes the likelihood
function when the dataset is known) is a normal distribution with
standard deviation $\sigma_i$ known from the error bars in the data set.
However, it is usually recommended to put in ``safety features'', in case the
data set contains any discrepant measurements, or in case the error bars in the
data set are underestimated. To achieve this, we used a student-$t$ distribution
instead of a normal distribution, with
scale parameter $\sqrt{\sigma_i^2 + s_0^2}$ and shape parameter $\nu$. The
parameter $s_0$ is an ``extra noise'' parameter that effectively increases the
size of the error bars, and the shape parameter $\nu$ allows for heavier tails
than a normal distribution. If $\nu$ is large, the student-$t$ distribution
is approximately a normal distribution, and if $\nu$ is small the noise
distribution has much heavier tails. For instance, when $\nu=1$ the student-$t$
distribution becomes a Cauchy distribution.

All of the model assumptions are specified in detail in
Table~\ref{tab:priors}. Cauchy distributions were used when there is a
preferred value, but since these have very heavy tails, the assumption is
quite fail safe. For example, the prior for $\mu_P$, the typical orbital period,
is centered around 1 year.

>>>>

The unknown parameters are:
\begin{eqnarray}
\left\{N, \boldsymbol{\alpha}, \{\boldsymbol{\psi}\}_{i=1}^N, y_0, \sigma_{\rm extra}, \nu\right\}
\end{eqnarray}
where $N$ is the number of planets,
$\boldsymbol{\alpha} = \{\mu_P, \sigma_P, \mu_A\}$ are the
hyperparameters used to define the prior for the properties of the planets,
and $\psi_i = \{P_i, A_i, \phi_i, e_i, \theta_i\}$
are the properties of planet $i$.
The parameter $y_0$ describes a DC offset in the data, and
$\sigma_{\rm extra}$ and $\nu$ are parameters of the noise distribution.

A standard assumption for the probability distribution of the data given the
parameters (known as the sampling distribution, which becomes the likelihood
function when the dataset is known) is a normal distribution with
standard deviation $\sigma_i$ known from the error bars in the data set.
However, it is usually recommended to put in ``safety features'', in case the
data set contains any discrepant measurements, or in case the error bars in the
data set are underestimated. To achieve this, we used a student-$t$ distribution
instead of a normal distribution, with
scale parameter $\sqrt{\sigma_i^2 + \sigma_{\rm extra}^2}$ and shape parameter $\nu$. The
parameter $\sigma_{\rm extra}$ is an ``extra noise'' parameter that effectively increases the
size of the error bars, and the shape parameter $\nu$ allows for heavier tails
than a normal distribution. If $\nu$ is large, the student-$t$ distribution
is approximately a normal distribution, and if $\nu$ is small the noise
distribution has much heavier tails. For instance, when $\nu=1$ the student-$t$
distribution becomes a Cauchy distribution.

All of the model assumptions are specified in detail in
Table~\ref{tab:priors}. Most of these priors were chosen to represent vague
prior knowledge, rather than the judgement of an informed expert on extrasolar
planets.

Cauchy distributions were used when there is a
preferred value, but since these have very heavy tails, the assumption is
quite fail safe. For example, the prior for $\mu_P$, the typical orbital period,
is centered around 1 year.
